---
title: "CS 378 Data Science Project1: ggplot/shiny"
output: html_document
---


##A look at natural disaster records 

####Our Data Sets

* Global earthquake records from the last 50 years (~570,000 rows)
* Global tsunami source data (~2,500 rows)
* Global tsunami runup data (~25,000 rows)
    + includes tsunamis from approximately the last two thousand years, but mostly from the last few hundred
* Significant volcanic eruptions data (~650 rows)
    + includes eruptions over a long period of time (thousands of years) but predominantly contains information from the 1900s


####This is how RStudio is configured:
```{r}
sessionInfo()
```

####Connecting to oracle and reading the data into RStudio

```{r}
Sys.setenv(JAVA_HOME="")
options(java.parameters="-Xmx2g")
library(rJava)

# Output Java version
.jinit()
print(.jcall("java/lang/System", "S", "getProperty", "java.version"))

# Load RJDBC library
library(RJDBC)

jdbcDriver <- JDBC(driverClass="oracle.jdbc.OracleDriver", classPath="C:/Program Files/Java/jdk1.7.0_67/ojdbc6.jar")

possibleError <- tryCatch(
  # Establishing the connection
  jdbcConnection <- dbConnect(jdbcDriver, "jdbc:oracle:thin:@128.83.138.158:1521:orcl", "C##cs347_mas5774", "orcl_mas5774"),
  error=function(e) e
)
if(!inherits(possibleError, "error")){
  # Reading in the data
  seismic_data <- dbGetQuery(jdbcConnection, "select * from seismic_data")
  tsunami_runup_data <- dbGetQuery(jdbcConnection, "select * from tsunami_runup_data")
  tsunami_source_data <- dbGetQuery(jdbcConnection, "select * from tsunami_source_data")
  volcano_data <- dbGetQuery(jdbcConnection, "select * from volcano_data")
  dbDisconnect(jdbcConnection)
}

```

####Some practice plots
```{r}

```

####Sources for the Data
[__ANSS (Advanced National Seismic System) Global Earthquake Catalog__](http://ncedc.org/anss/catalog-search.html)

* hosted by the Northern California Earthquake Data Center
* search parameters used:
    + catalog = ANSS
    + start_time = 1964/01/01,00:00:00
    + end_time = 2014/09/12,22:13:44
        + covers the last 50 years, anything larger would have been unmanageable/unrealistic at this point
    + minimum_magnitude = 3.0
        + 3.0 was chosen as a minimum because that is generally the minimum threshold for people to notice an earthquake
    + maximum_magnitude = 10
    + event_type = E
    
[__National Geophysical Data Center / World Data Service (NGDC/WDS): Significant Volcanic Eruptions Database__](http://www.ngdc.noaa.gov/nndc/servlet/ShowDatasets?dataset=102557&search_look=50&display_look=50)

* hosted by the National Geophysical Data Center of the National Oceanic and Atmospheric Administration
* search parameters were all left blank (so we used the entire data set, unfiltered)

[__National Geophysical Data Center / World Data Service (NGDC/WDS): Global Historical Tsunami Database__](http://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=70&d=7)

* hosted by the National Geophysical Data Center of the National Oceanic and Atmospheric Administration
* search parameters were all left blank (again using the entire data set, unfiltered)
* the url above links to the 'event' information, you can get to the 'runup' information [__here__](http://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=167&d=166)